---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

## Visualize Benchmark Results

```{r}
bmrdt_measured <- read.csv("bmrdt_measured.csv")
```

### Main results

We are interested in how well all calibration methods and strategies
improved the calibration of the base learners compared to the
uncalibrated base learner:

```{r}
library(data.table)
library(ggplot2)

ggplot(bmrdt_measured, aes(x = base_learner, y = classif.ece, fill = calibration_method)) +
  geom_boxplot() +
  theme_minimal(base_size = 14) +
  labs(
    x = "Base Learner",
    y = "Classification ECE",
    fill = "Calibration Method",
    title = "ECE per Base Learner and Calibration Method"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(bmrdt_measured, aes(x = base_learner, y = classif.ece, fill = calibration_method)) +
  geom_boxplot() +
  facet_wrap(~ calibration_strategy, scales = "fixed") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Base Learner",
    y = "Classification ECE",
    fill = "Calibration Method",
    title = "ECE per Base Learner and Calibration Method (by Calibration Strategy)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Naive Bayes is per se poorly calibrated, whereas Random Forest is
already well calibrated before applying any of the calibration methods
and strategies.

There is no real difference visible between `oof` and `perfold`
calibration: Both improve Naive Bayes Calibration in a similar manner,
both do not quite reduce the variance of `ece` compared the uncalibrated
base learner.

When only looking at the calibration method (first plot) all methods
seem to work equally well for Naive Bayes and Random Forest. For
Decision Tree, `platt` slightly outperforms the other methods.

### Deep Dive Calibration Strategy

```{r}
ggplot(bmrdt_measured, aes(x = base_learner, y = classif.ece, fill = calibration_strategy)) +
  geom_boxplot() +
  theme_minimal(base_size = 14) +
  labs(
    x = "Base Learner",
    y = "Classification ECE",
    fill = "Calibration Strategy",
    title = "ECE per Base Learner and Calibration Strategy"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Calibration only really helps for Naive Bayes. There is no difference
visible between the strategies.

### Deep Dive Differences between Benchmark Tasks

```{r}
library(data.table)
ggplot(
  subset(bmrdt_measured, calibration_strategy == "none"),
  aes(x = task_id, y = classif.ece, color = base_learner)
) +
  geom_boxplot() +
  theme_minimal(base_size = 14) +
  labs(
    x = "Task ID",
    y = "Classification ECE",
    fill = "Base Learner",
    title = "ECE per Task (No Calibration Strategy)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Aggregate ECE per task_id and base_learner where calibration_strategy is none
library(dplyr)

plotdt <- bmrdt_measured %>%
  filter(calibration_strategy == "none") %>%
  group_by(task_id, base_learner) %>%
  summarise(mean_ece = mean(classif.ece, na.rm = TRUE), .groups = "drop")


# Scatter plot of averaged ECE
ggplot(plotdt, aes(x = task_id, y = mean_ece, color = base_learner, group = base_learner)) +
  geom_point(size = 3) +
  geom_line() +  # optional: connect points per base_learner
  theme_minimal(base_size = 14) +
  labs(
    x = "Task ID",
    y = "Mean Classification ECE",
    color = "Base Learner",
    title = "Average ECE per Task and Base Learner (No Calibration Strategy)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Naive Bayes is really calibrated worst for most, but also Decision Tree
seems similar except for `pc3`

```{r}
# Aggregate mean ECE per task_id, base_learner, and calibration_strategy
plotdt <- bmrdt_measured %>%
  group_by(task_id, base_learner, calibration_strategy) %>%
  summarise(mean_ece = mean(classif.ece, na.rm = TRUE), .groups = "drop")


# Scatter plot with facets for each base learner
ggplot(plotdt, aes(x = task_id, y = mean_ece, color = calibration_strategy, group = calibration_strategy)) +
  geom_point(size = 3) +
  geom_line() +
  facet_wrap(~ base_learner, scales = "free_x") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Task ID",
    y = "Mean Classification ECE",
    color = "Calibration Strategy",
    title = "Average ECE per Task, by Calibration Strategy, per Base Learner"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

similarity over strategies

```{r}
# Aggregate mean ECE per task_id, base_learner, calibration_strategy, and calibration_method
plotdt <- bmrdt_measured %>%
  group_by(task_id, base_learner, calibration_strategy, calibration_method) %>%
  summarise(mean_ece = mean(classif.ece, na.rm = TRUE), .groups = "drop")


# Create a combined factor for coloring
plotdt <- plotdt %>%
  mutate(strategy_method = paste(calibration_strategy, calibration_method, sep = "_"))


# Scatter plot with facets per base learner
ggplot(plotdt, aes(x = task_id, y = mean_ece, color = strategy_method, group = strategy_method)) +
  geom_point(size = 3) +
  geom_line() +
  facet_wrap(~ base_learner, scales = "free_x") +
  theme_minimal(base_size = 14) +
  labs(
    x = "Task ID",
    y = "Mean Classification ECE",
    color = "Strategy_Method",
    title = "Average ECE per Task by Calibration Strategy + Method per Base Learner"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
