---
title: "mlr3calibration"
author: "Adrian Glos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlr3calibration}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
header-includes:
  - \pagenumbering{gobble}
---

# Introduction

`mlr3calibration` extends `mlr3` with probability calibration workflows,
calibration measures, and reliability plots.
Build calibration workflows with `po()`, obtain calibrators with `clb()`,
evaluate with `msr()`, and inspect reliability with `calibration_plot()`.

Two recommended workflows:

1. calibrate a fixed learner or a full graph/pipeline,
2. jointly tune the calibrated object with an outer `AutoTuner`.

`tune_then_calibrate()` is a sequential one-call alternative and is not the recommended tuning workflow.

Binary classification is the main supported path.
Multiclass support is available through calibrators that declare the `multiclass` property.

Built-in calibrators:

- Platt scaling through `clb("platt")` (default: Platt soft targets as in the original paper / sklearn;
  set `label_smoothing = FALSE` for hard-label logistic calibration as in probably;
  set `input = "logit"` to follow Silva Filho et al. on probability inputs)
- Beta calibration through `clb("beta")`
- Isotonic regression through `clb("isotonic")`

# Installation

Install from GitHub:

```{r installation, eval=FALSE}
#remotes::install_github("giuseppec/mlr3calibration")
```

```{r libraries, message=FALSE, warning=FALSE}
library(mlr3)
library(mlr3calibration)
library(mlr3learners)
library(mlr3pipelines)
library(mlr3tuning)

has_rpart = mlr3misc::require_namespaces("rpart", quietly = TRUE)

set.seed(1)
```

# Setup

```{r task, eval=has_rpart}
data("Sonar", package = "mlbench")
task = as_task_classif(Sonar, target = "Class", positive = "M")
splits = partition(task)
task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)

measures = msrs(c(
  "classif.bbrier",
  "classif.ece",
  "classif.ici",
  "classif.hltest",
  "classif.cox_intercept",
  "classif.cox_slope",
  "classif.spiegelhaltersz"
))
```

# Workflow 1: calibrate a fixed learner or full graph

Wrap a `Learner`, `GraphLearner`, or bare `Graph` in `po("calibrate", ...)`.
Put preprocessing that should participate in calibration resampling inside
`learner = ...`; prefer wrapping the whole graph over
`po("scale") %>>% po("calibrate", learner = ...)`.

```{r workflow-fixed-graph, eval=has_rpart}
base_graph = po("scale") %>>%
  po("learner", learner = lrn("classif.rpart", predict_type = "prob"))

learner_uncal = as_learner(base_graph$clone(deep = TRUE))
learner_cal = as_learner(po(
  "calibrate",
  learner = base_graph$clone(deep = TRUE),
  calibrator = clb("platt"),
  strategy = "oof",
  resampling = rsmp("cv", folds = 3)
))

learner_uncal$id = "uncalibrated"
learner_cal$id = "calibrated"
```

```{r workflow-fixed-graph-train, warning=FALSE, eval=has_rpart}
learner_uncal$train(task_train)
learner_cal$train(task_train)

prediction_uncal = learner_uncal$predict(task_test)
prediction_cal = learner_cal$predict(task_test)
```

# Measures and plots

```{r measures, eval=has_rpart}
prediction_uncal$score(measures)
prediction_cal$score(measures)
```

```{r calibration-plot, fig.height=5, fig.width=9, warning=FALSE, eval=has_rpart}
calibration_plot(list(learner_uncal, learner_cal), task_test, smooth = TRUE)
calibration_plot(predictions = list(
  uncalibrated = prediction_uncal,
  calibrated = prediction_cal
), smooth = TRUE)
```

# Calibration-error measures

Formulas, edge conventions, and sources live on the measure help pages
(`?MeasureClassifConfidenceECE`, `?MeasureClassifSCE`, `?MeasureClassifACE`,
`?MeasureClassifTACE`, `?MeasureClassifTopLabelECE`, `?MeasureClassifECE`).

```{r calibration-measures}
multiclass_measures = msrs(c(
  "classif.conf_ece",
  "classif.sce",
  "classif.ace",
  "classif.tace",
  "classif.tl_ece"
))
multiclass_measures
```

| Dictionary key | Meaning | Source |
|---|---|---|
| `classif.ece` | Binary positive-class ECE (fixed-width bins) | Kull et al. (2019), Eq. 4 |
| `classif.conf_ece` | Confidence ECE (max prob vs correctness) | Guo et al. (2017), Eq. 3 |
| `classif.sce` | Classwise ECE / static calibration error | Nixon et al. (2019); Kull et al. (2019), Eq. 4 |
| `classif.ace` | Classwise adaptive calibration error | Nixon et al. (2019) |
| `classif.tace` | Thresholded ACE | Nixon et al. (2019) |
| `classif.tl_ece` | Top-label ECE | Gupta and Ramdas (2022), Eqs. 2 and 4 |

Default `bins` / `ranges` are package defaults, not part of the published definition.

What each measure conditions on:

- `classif.conf_ece`: maximum probability and whether the predicted label is correct (labels pooled).
- `classif.tl_ece`: same, but also conditional on the predicted label (never-predicted classes contribute nothing).
- `classif.sce`: every class probability on all one-vs-rest observations, fixed-width bins, equal class average.
- `classif.ace`: same classwise target as `classif.sce`, adaptive equal-count ranges.
- `classif.tace`: `classif.ace` restricted to class probabilities strictly above `epsilon`.
- `classif.ece`: binary positive-class workflow.

Fixed-width bins use right-closed intervals; probability zero goes in the first bin; empty bins have zero weight and do not renormalize.
Adaptive ranges sort by probability then row ID, and when `N` is not divisible by `R` the first `N mod R` ranges get one extra observation.
Every class needs at least `R` retained observations; `classif.tace` errors if a class has too few values above the threshold.
Classwise measures are macro-averaged (not prevalence-weighted); tiny non-target probabilities can make a classwise scalar look small, which is what `classif.tace` addresses by thresholding.
None of these alone is full multiclass calibration; interpret them with a proper score (log loss or Brier) and, where useful, reliability plots.

Not provided: micro ECE over all class-probability pairs, prevalence-weighted classwise ECE, macro average of per-predicted-label top-label ECE, multivariate multiclass ICI, or multiclass Hosmer-Lemeshow / Cox / Spiegelhalter measures.

# Workflow 2: jointly tune the calibrated object

Wrap the calibrated learner in an outer `AutoTuner`.
To tune over several calibrators, use `clb("selector")` with
`calibrator_selector_search_space()`.

```{r workflow-outer-autotuner, warning=FALSE, eval=has_rpart}
learner_joint = as_learner(po(
  "calibrate",
  learner = lrn("classif.rpart", predict_type = "prob"),
  calibrator = clb("selector", choices = c("platt", "beta")),
  strategy = "oof",
  resampling = rsmp("cv", folds = 2)
))

search_space = paradox::ParamSetCollection$new(c(
  list(paradox::ps(
    "calibrate.classif.rpart.cp" = paradox::p_dbl(0.001, 0.1)
  )),
  list(calibrator_selector_search_space(choices = c("platt", "beta")))
))

autotuner_joint = AutoTuner$new(
  learner = learner_joint,
  resampling = rsmp("cv", folds = 2),
  measure = msr("classif.bbrier"),
  search_space = search_space,
  terminator = bbotk::trm("evals", n_evals = 4),
  tuner = tnr("grid_search", resolution = 2)
)

suppressWarnings(autotuner_joint$train(task_train))
autotuner_joint$predict(task_test)$score(msr("classif.bbrier"))
```

# Convenience helper: tune_then_calibrate

`tune_then_calibrate()` tunes on the full training task, then calibrates that fixed tuned learner.
Hyperparameters are not jointly tuned with calibration; prefer an outer `AutoTuner` around `po("calibrate", ...)` instead.

```{r tune-then-calibrate, warning=FALSE, eval=has_rpart}
inner_at = auto_tuner(
  tuner = tnr("grid_search", resolution = 2),
  learner = lrn(
    "classif.rpart",
    predict_type = "prob",
    cp = paradox::to_tune(0.001, 0.1)
  ),
  resampling = rsmp("cv", folds = 2),
  measure = msr("classif.bbrier"),
  term_evals = 4
)

learner_seq = tune_then_calibrate(
  learner = inner_at,
  calibrator = clb("platt"),
  strategy = "oof",
  resampling = rsmp("cv", folds = 2)
)

learner_seq$train(task_train)
learner_seq$predict(task_test)$score(msr("classif.bbrier"))
```

# Reusing a precomputed resample result

Compare several calibrators on the same out-of-fold predictions by computing the resampling once:

```{r rr, warning=FALSE, eval=has_rpart}
rr = resample(
  task_train,
  lrn("classif.rpart", predict_type = "prob"),
  rsmp("cv", folds = 3),
  store_models = TRUE
)

learner_platt = as_learner(po(
  "calibrate",
  rr = rr,
  calibrator = clb("platt"),
  strategy = "oof"
))

learner_beta = as_learner(po(
  "calibrate",
  rr = rr,
  calibrator = clb("beta", parameters = "ab"),
  strategy = "oof"
))

scores = data.frame(
  platt = learner_platt$train(task_train)$predict(task_test)$score(msr("classif.ece")),
  beta = learner_beta$train(task_train)$predict(task_test)$score(msr("classif.ece"))
)

scores
```

The precomputed result must be from the same calibration task and contain stored probability learners with complete out-of-fold test predictions.
`learner`, `resampling`, and learner hyperparameters cannot be supplied in this mode.
With `strategy = "oof"`, a single uncalibrated model is still refitted on the full training task using the stored learners' configuration.

# Multiclass support

Use `clb("ovr")` for one-vs-rest wrapping.
Direct `clb("platt")` is binary-only; multiclass Platt is `clb("ovr")`.

```{r multiclass, warning=FALSE, eval=has_rpart}
task_iris = as_task_classif(iris, target = "Species")
splits_iris = partition(task_iris)
task_iris_train = task_iris$clone()$filter(splits_iris$train)
task_iris_test = task_iris$clone()$filter(splits_iris$test)

learner_multiclass = as_learner(po(
  "calibrate",
  learner = lrn("classif.rpart", predict_type = "prob"),
  calibrator = clb("ovr", calibrator = clb("isotonic")),
  strategy = "oof",
  resampling = rsmp("cv", folds = 3)
))

learner_multiclass$train(task_iris_train)
head(learner_multiclass$predict(task_iris_test)$prob)
```

# Writing a custom calibrator

Subclass [Calibrator], register with `register_calibrator()`, and construct with `clb()`.
The example below shows the extension hooks only; it is not a recommended calibrator.

```{r custom-calibrator, warning=FALSE, eval=has_rpart}
TemperatureCalibrator = R6::R6Class(
  "TemperatureCalibrator",
  inherit = Calibrator,
  public = list(
    initialize = function(temperature = 1) {
      param_set = paradox::ps(
        temperature = paradox::p_dbl(lower = 0.1, upper = 5, default = 1, tags = "train")
      )
      super$initialize(
        id = "temperature_example",
        param_set = param_set,
        properties = c("twoclass", "multiclass"),
        label = "Temperature example calibrator"
      )
      self$param_set$values$temperature = temperature
    }
  ),
  private = list(
    .temperature = NULL,
    .reset = function() {
      private$.temperature = NULL
    },
    .train = function(truth, prob, class_names) {
      temperature = self$param_set$values$temperature
      if (is.null(temperature)) {
        temperature = 1
      }
      private$.temperature = temperature
    },
    .predict = function(prob, class_names) {
      logits = log(pmax(prob, .Machine$double.eps))
      exp(logits / private$.temperature)
    }
  )
)

register_calibrator("temperature_example", TemperatureCalibrator)
calibrator = clb("temperature_example", temperature = 1.5)

learner_custom = as_learner(po(
  "calibrate",
  learner = lrn("classif.rpart", predict_type = "prob"),
  calibrator = calibrator,
  strategy = "oof",
  resampling = rsmp("cv", folds = 2)
))

learner_custom$train(task_train)
print(learner_custom$predict(task_test)$score(msr("classif.bbrier")))
mlr3_calibrators$remove("temperature_example")
```

Discover keys with `mlr3_calibrators$keys()`.
Calibrators with required constructor arguments can be registered and built via `clb(key, ...)`, but they can enter `CalibratorSelector` only when `clb(key)` works without arguments.
