---
title: "mlr3calibration"
author: "Adrian Glos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlr3calibration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
  - \pagenumbering{gobble}
---

# General
This package is a flexible calibration extension for the mlr3 framework in R. It includes the ability to calibrate objects of type learner, so that they can be trained or predicted as usual. It allows to choose between Platt scaling, Beta calibration and isotonic calibration. Own Calibrators can be included as well. In addition, a calibrated learner can either be easily integrated into pipelines or entire pipelines can be calibrated. The calibrated learner can also be initialized by a base learner with tune token, allowing parameter tuning for the calibrated learner.This package also provides the ability to plot the reliability curve for one or more trained learners and a task. Furthermore, this extension provides several calibration measures: the expected calibration error, the integrated calibration index, HL-test, Cox Slope and Intercept and Spiegelhalter's Z statistics to measure calibration.

# Installation

You can install the mlr3calibration package from GitHub with the following code:

```{r installation, message=FALSE, warning=FALSE}
#remotes::install_github("giuseppec/mlr3calibration")
# library(mlr3calibration)
# do devtools::load_all()
library(mlr3verse)
```

# Calibration

To use the mlr3calibration package, we first need a binary classification task. mlr3 does not support multiclass calibration yet. One possibility to do multiclass calibration is to calibrate several one-vs.-rest binary tasks.
```{r task}
# Load a binary classification task
set.seed(1)
data("Sonar", package = "mlbench")
task = as_task_classif(Sonar, target = "Class", positive = "M")
splits = partition(task)
task_train = task$clone()$filter(splits$train)
task_test = task$clone()$filter(splits$test)
```

To calibrate a learner you need an uncalibrated learner, a resampling strategy and a calibration method (platt, beta or isotonic). To prevent overfitting, calibration can be performed using cross-validation. The dataset is divided into k folds, and each fold is used for calibration once, while the other (k-1) folds are used for training the classifier. If, for example, a holdout resampling strategy is selected, then no cross-validated calibration is performed, but the base learner is trained on the training split and the calibrator on the holdout.
There are several strategies when to train the calibrator with cross validation:
- OOF (Out-Of-Fold Calibration) `calib.oof`:
Train calibration on out-of-fold predictions from cross-validation. Uses all data without leakage.
- perFold Calibration `calib.perfold`:
Each fold gets its own calibrator trained only on that fold’s validation predictions. Multiple calibrators exist.
- OOFTuneFirst `calib.tunefirst.oof`:
First, tune the base learners hyperparamerters, then calibrate on OOF predictions (global view) using the tuned learner. Needs an AutoTuner as learner object.
- PerFoldTuneFirst `calib.tunefirst.perfold`:
First, tune the base learners hyperparamerters, then calibrate on within each fold using the tuned learner. Needs an AutoTuner as learner object.

```{r calibration, warnings = FALSE}
# Initialize the uncalibrated learner
learner_uncal <- lrn("classif.xgboost", nrounds = 50, predict_type = "prob")

# Initialize the calibrated learner
rsmp <- rsmp("cv", folds = 5)

learner_cal <- as_learner(calibPipeOp("calib.oof",
                                      learner = learner_uncal,
                                      method = "beta",
                                      parameters = "ab",
                                      resampling = rsmp))

# Set ID's for the learners
learner_uncal$id <- "Uncalibrated Learner"
learner_cal$id <- "Calibrated Learner"
```
The calibrated learner can be trained in the same way as the base learner.
```{r train, results = "hide", warning=FALSE}
# Train the learners
learner_uncal$train(task_train)
learner_cal$train(task_train)
```

It is also possible to implement an own calibrator as a R6 class. This calibrator can then be added to the calibrator dictionary of this library. Here, we create a calibrator `Platt2` that is equal to `Platt` Calibration.

```{r}
# write the calibrator class that inherits from LearnerClassif
CalibratorPlatt2 = R6::R6Class("CalibratorPlatt2",
                               inherit = LearnerClassif,

                               public = list(

                                 #' @description
                                 #' Creates a new instance of this [R6][R6::R6Class] class.
                                 initialize = function() {

                                   ps = paradox::ps()

                                   super$initialize(
                                     id = "platt2",
                                     param_set = ps,
                                     predict_types = c("response", "prob"),
                                     feature_types = c("logical", "integer", "numeric", "character", "factor", "ordered"),
                                     properties = c("weights", "twoclass", "offset"),
                                     packages = c("mlr3learners", "stats"),
                                     label = "platt",
                                   )

                                   private$.calibrator = mlr3::lrn("classif.log_reg", predict_type = "prob")

                                 }

                               ),
                               private = list(
                                 .calibrator = NULL,

                                 .train = function(task) {
                                   # For example, fit a logistic regression model
                                   private$.calibrator$train(task)
                                 },

                                 .predict = function(task) {
                                   pred_calibrated = private$.calibrator$predict(task)
                                   return(pred_calibrated)
                                 }
                               )
)

# register the new calibrator
register_calibrator("platt2", CalibratorPlatt2)

# use the new calibrator
# Initialize the uncalibrated learner
learner_uncal <- lrn("classif.xgboost", nrounds = 50, predict_type = "prob")

# Initialize the calibrated learner
rsmp <- rsmp("cv", folds = 5)

learner_cal <- as_learner(calibPipeOp("calib.oof",
                                      learner = learner_uncal,
                                      method = "platt2",
                                      resampling = rsmp))

# Set ID's for the learners
learner_uncal$id <- "Uncalibrated Learner"
learner_cal$id <- "Calibrated Learner"

# Train the learners
learner_uncal$train(task_train)
learner_cal$train(task_train)
```

# Calibration Pipelines

We have already seen that a calibrator can be applied to an uncalibrator learner using a pipeline with a calibration strategy (Out-Of-Fold, In-Fold, Tune-First-OOF, Tune-First-In-Fold) and a resampling.
It is possible to implement an own Pipeline as well.

Imagine we want to do OOF Calibration using Hold-Out-Split as a default. We could write our own pipeline for this and add it to the Pipeline dictionary of this library:

```{r}
PipeOpCalibrationOOFHoldOut <- R6::R6Class(
  "PipeOpCalibrationOOFHoldOut",
  inherit = mlr3pipelines::PipeOp,

  public = list(
    learner = NULL,
    method = NULL,
    resampling = NULL,
    learners = NULL,
    calibrator = NULL,
    parameters = NULL,
    rr = NULL,

    #' @description
    #' Creates a new `PipeOpCalibrationPerFold` object.
    #' @param learner Base learner to be calibrated. predict_type has to be `"prob"`.
    #' @param rr [`ResampleResult`][mlr3::ResampleResult]\cr
    #'   Optional pre-computed result; skips the internal resampling step.
    #' @param method Calibration method to use. One of `"platt"`, `"isotonic"`, or `"beta"`. Default is `"platt"`.
    #' @param resampling Resampling strategy for cross-validation. Default is `rsmp("cv", folds = 5)`.
    #' @param parameters Parameters for beta calibration. Default is `"abm"`.
    #' @param param_vals param_vals, copied from base learner
    initialize = function(#id = paste0(self$learner$id, ".calibrated_oof_", method),
      learner,
      method = "platt",
      resampling = rsmp("cv", folds = 2),
      rr = NULL,
      parameters = NULL,
      param_vals = list()) {
      if (!is.null(rr)) self$rr <- rr
      self$learner = learner
      self$method = method
      self$resampling = resampling
      self$learners = list()
      self$parameters = parameters
      super$initialize(id = self$learner$base_learner()$id, 
        param_set = alist(self$learner$param_set),
        param_vals = param_vals,
        input = data.table(name = "input", train = "Task",
          predict = "Task"),
        output = data.table(name = "output", train = "NULL",
          predict = "PredictionClassif"),
      )
    }
  ),
  active = list(
    predict_type = function(val) {
      if (!missing(val)) {
        assert_subset(val, names(mlr_reflections$learner_predict_types[[self$learner$task_type]]))
        self$learner$predict_type = val
      }
      self$learner$predict_type
    }
  ),
  private = list(
    .train = function(inputs) {
      on.exit(lgr::get_logger("mlr3")$set_threshold("info"))

      lgr::get_logger("mlr3")$set_threshold("warn")

      task = inputs[[1]]
      positive = task$positive

      rr <- if (is.null(self$rr)) {
        resample(task, self$learner, self$resampling, store_models = TRUE)
      } else self$rr
      self$learners = rr$learners

      preds = rr$predictions(predict_sets = "test")
      preds = lapply(preds, as.data.table)
      pred_data = data.table::rbindlist(preds, use.names = TRUE)

      calibration_data = data.table(truth = pred_data$truth,
        response = with(pred_data, get(paste0("prob.", positive))))
      colnames(calibration_data) = c("truth", "response")
      calibration_data$response = as.numeric(calibration_data$response)
      task_for_calibrator = as_task_classif(calibration_data,
                                            target = "truth",
                                            positive = positive,
                                            id = "task_cal")

      if(is.null(self$parameters)){
        calibrator <- clb(self$method)
      }else{
        calibrator <- clb(self$method, self$parameters)
      }
      self$calibrator = calibrator
      self$calibrator$train(task_for_calibrator)

      return(list(NULL))
    },

    .predict = function(inputs) {
      task = inputs[[1]]
      positive = task$positive
      predictions = list()
      for (learner_index in seq_along(self$learners)) {
        learner = self$learners[[learner_index]]
        pred = learner$predict(task)
        predictions[[length(predictions) + 1]] = as.data.table(pred)
      }
      response = rowMeans(sapply(predictions, function(x) as.numeric(x$response)))
      response = ifelse(response < 1.5, positive, task$negative)
      prob = as.matrix(data.frame(
        rowMeans(sapply(predictions, function(x) with(x, get(paste0("prob.", positive))))),
        rowMeans(sapply(predictions, function(x) with(x, get(paste0("prob.", task$negative)))))
      ))
      colnames(prob) = c(task$positive, task$negative)
      pred_uncal = PredictionClassif$new(
        task = task,
        row_ids = task$row_ids,
        truth = task$truth(),
        prob = prob,
        response = response
      )
      pred_uncal = as.data.table(pred_uncal)

      calibration_data = data.table(truth = pred_uncal$truth,
                                    response = with(pred_uncal, get(paste0("prob.", positive))))
      colnames(calibration_data) = c("truth", "response")
      calibration_data$response = as.numeric(calibration_data$response)
      task_for_calibrator = as_task_classif(calibration_data,
                                            target = "truth",
                                            positive = positive,
                                            id = "task_cal")

      pred_calibrated = self$calibrator$predict(task_for_calibrator)

      return(list(pred_calibrated))
    },

    .additional_phash_input = function() {
      list(self$learner$hash, self$calibration_ratio)
    }
  )
)

# Register the new PipeOp
register_calibration_pipeop("calib.oof.holdout", PipeOpCalibrationOOFHoldOut)
```

Now, use the newly created pipeline.


```{r}
# Initialize the uncalibrated learner
learner_uncal <- lrn("classif.xgboost", nrounds = 50, predict_type = "prob")


# Initialize the calibrated learner
rsmp <- rsmp("cv", folds = 5)

learner_cal <- as_learner(calibPipeOp("calib.oof.holdout",
                                      learner = learner_uncal,
                                      method = "platt"))

# Set ID's for the learners
learner_uncal$id <- "Uncalibrated Learner"
learner_cal$id <- "Calibrated Learner"

# Train the learners
learner_uncal$train(task_train)
learner_cal$train(task_train)
```



# Calibration Measures

To measure the calibration, this package provides the Expected Calibration Error (ECE) and the Integrated Calibration Index (ICI). The ECE is a measure of the difference between the predicted probabilities and the true outcomes. The ICI is a weighted average of the absolute differences between the calibration curve and the diagonal perfectly calibrated line.

```{r measure}
# Predict the Learners
preds_uncal <- learner_uncal$predict(task_test)
preds_cal <- learner_cal$predict(task_test)
# Calculate the Expected Calibration Error (ECE)
ece_uncal <-  preds_uncal$score(ece$new())
ece_cal <-  preds_cal$score(ece$new())
# Uncalibrated ECE
ece_uncal
# Calibrated ECE
ece_cal
```

# Raliability Curve

To visualize how well the model is calibrated, this package provides a calibration plot. The calibration plot shows the relationship between the predicted probabilities and the true outcomes. The plot is divided into bins, and within each bin, the mean predicted probability and the mean observed outcome are calculated. The calibration plot can be smoothed using LOESS.

```{r calibrationplot, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
# List the Learners you want to plot
lrns = list(learner_uncal, learner_cal)

# Plot the reliability curve
library(ggplot2)
calibrationplot(lrns, task_test, smooth = TRUE)
```

# More Calibration Measures

## HL-Test

Tests the goodness of calibration for a binary classifier by comparing expected and observed frequencies in data bins.
A small p-value indicates that the model is poorly calibrated to the data.

```{r}
preds_uncal = learner_uncal$predict(task_test)
preds_cal = learner_cal$predict(task_test)
preds_uncal$score(hltest$new())
preds_cal$score(hltest$new())
```
## Cox Test
A logistic regression model with log-odds as the independent variable and the outcome as the dependent variable is fitted. Perfect calibration is indicated by a slope of 1 and intercept of 0. Slope >1 indicates overconfidence at high probabilities and underconfidence at low probabilities, while a slope <1 indicates the opposite. A positive intercept indicates general overconfidence.

```{r}
preds_uncal = learner_uncal$predict(task_test)
preds_cal = learner_cal$predict(task_test)
preds_uncal$score(cox_slope$new())
preds_cal$score(cox_slope$new())
```

## Spiegelhalters Z Statistics

Spiegelhalter's test compares predicted probabilities to observed outcomes and tests if the model is well calibrated.
A small p-value (<0.05) indicates bad calibration.

```{r}
preds_uncal = learner_uncal$predict(task_test)
preds_cal = learner_cal$predict(task_test)
preds_uncal$score(spiegelhaltersz$new())
preds_cal$score(spiegelhaltersz$new())
```

## Own Calibration Measures

It is also possible to implement own calibration measures. For example, we can implement the Maximum Calibration Error (MCE):

Each measure is a R6 class that inherits from `mlr3::MeasureClassif` and implements a `score` function that gives back the measure value for some prediction.

```{r}
mce = R6::R6Class("mce",
                  inherit = mlr3::MeasureClassif,
                  public = list(
                    #' @description
                    #' Creates a new `mce` object.
                    initialize = function() {
                      super$initialize(
                        id = "classif.mce",
                        properties = character(),
                        predict_type = "prob",
                        range = c(0, 1),
                        minimize = TRUE
                      )
                    }
                  ),
                  private = list(
                    .score = function(prediction, ...) {
                      actual = ifelse(prediction$truth == colnames(prediction$prob)[1], 1, 0)
                      predicted = prediction$prob[, 1]
                      n_bins = 10
                      # Create bins
                      bins <- cut(predicted, breaks = seq(0, 1, length.out = n_bins + 1), include.lowest = TRUE)
                      
                      # For each bin: compute accuracy (mean of y_true) and confidence (mean of y_prob)
                      bin_summary <- aggregate(
                        data.frame(actual = actual, predicted = predicted), 
                        by = list(bin = bins), 
                        FUN = mean
                      )
                      
                      # Calibration error per bin
                      bin_summary$calib_error <- abs(bin_summary$actual - bin_summary$predicted)
                      
                      # Maximum Calibration Error
                      mce_value <- max(bin_summary$calib_error, na.rm = TRUE)
                      
                      return(mce_value)
                    }
                  )
)
```

The new measure has to be added to the mlr3 measures dictinary to be accessible via key String.

```{r}
mlr3::mlr_measures$add("classif.mce", mce)
```


Now, let us try out the new measure on the existing calibrated and uncalibrated learner.

```{r}
preds_uncal = learner_uncal$predict(task_test)
preds_cal = learner_cal$predict(task_test)
preds_uncal$score(mce$new())
preds_cal$score(mce$new())
```



# Calibration with resample Object

If you want to compare the three calibration approaches with each other, it can be time-consuming to train each calibration approach using cross validation. In these cases, you can train an rr object, store the models and pass it to the PipeOpCalibrationPerFold. This means that the base learner has to be trained once in advance, and only the calibrators have to be trained in the PipeOpCalibrationPerFold using the holdout folds.

```{r Calibration via rr Object, results = "hide", warning=FALSE}
# Initialize base learner
learner_xg <- lrn("classif.xgboost", predict_type = "prob")

# Initialize resampling strategy
rsmp <- rsmp("cv", folds = 5)

# Create rr object
rr <- resample(task_train, learner_xg, rsmp, store_models = TRUE)

# Initialze the calibrated learners
learner_calibrated_platt <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "platt"))

learner_calibrated_beta <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "beta"))

learner_calibrated_isotonic <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "isotonic"))

# Train the calibrated learners
learner_calibrated_platt$train(task_train)
learner_calibrated_beta$train(task_train)
learner_calibrated_isotonic$train(task_train)
```

# Pipelines

It is also possible to calibrate a mlr3 pipeline, or to include an Calibrated Learner in a pipeline.

## Calibrate a Pipeline

```{r Calibrate a Pipeline, results = "hide", warning=FALSE}
# Calibrate a pipeline
pipeline <- as_learner(po("imputemean") %>>% lrn("classif.ranger", 
                                                 predict_type = "prob"))

learner_cal <- as_learner(calibPipeOp("calib.perfold",
                        learner = pipeline,
                        method = "platt",
                        resampling = rsmp("cv", folds = 5)))

learner_cal$train(task_train)
```

## Include Calibrated Learner in a Pipeline

```{r Include Calibrated Learner in Pipeline, results = "hide", warning=FALSE}
# Include Calibrated learner in a pipeline

learner_cal <- as_learner(calibPipeOp("calib.perfold",
                        learner = lrn("classif.ranger", 
                                      predict_type="prob"),
                        method = "platt",
                        resampling = rsmp("cv", folds = 5)))

pipeline <- as_learner(po("imputemean") %>>% learner_cal)

pipeline$train(task_train)
```

# Tuning

You can either pass a tuned base learner to the calibrator or tuning your base learner within the calibrator.

```{r Tuning, results = "hide", warning=FALSE}
library(mlr3tuning)

# Include Calibrated learner in Pipeline
learner_uncal <- lrn("classif.ranger",
                     predict_type = "prob",
                     mtry = to_tune(1, 10),
                     num.trees = to_tune(100, 500))


learner_cal <- as_learner(calibPipeOp("calib.perfold",
                        learner = learner_uncal,
                        method = "platt",
                        resampling = rsmp("cv", folds = 5)))


at = auto_tuner(
  tuner = tnr("random_search"),
  learner = learner_cal,
  resampling = rsmp ("holdout"),
  measure = msr("classif.bbrier"),
  term_evals = 4)

at$train(task_train)
```

# Sample Use Case

Our goal for this exercise sheet is to learn the basics of model calibration for supervised classification with `mlr3calibration`.
In a calibrated model, the predicted probability for an input feature vector can be interpreted as the true likelihood of the outcome belonging to the positive class, meaning that among all instances assigned a probability of $p$, approximately $p\%$ will belong to the positive class.

## Required Packages

We will use `mlr3` for machine learning, and `mlr3calibration` specifically for calibration:

```{r}
#library(mlr3calibration)
library(mlr3verse)
library(mlr3learners)

set.seed(12345)
```

## Data: predicting cell segmentation quality

The `modeldata` package contains a data set called `cells`. Initially distributed by [Hill and Haney (2007)](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-8-340), they showed how to create models that predict the _quality_ of the image analysis of cells. The outcome has two levels: `"PS"` (for poorly segmented images) or `"WS"` (well-segmented). There are 56 image features that can be used to build a classifier. 

Let's load the data and remove an unwanted column: 

```{r}
library(modeldata)
data(cells, package = "modeldata")
cells$case <- NULL
```

## Uncalibrated Naive Bayes Learner

We'll show the utility of calibration tools by using a type of model that, in this instance, is likely to produce a poorly calibrated model. The Naive Bayes classifier is a well-established model that assumes that the predictors are statistically _independent_ of one another (to simplify the calculations).  While that is certainly not the case for this data, the model can be effective at discriminating between the classes. Unfortunately, when there are many predictors in the model, it has a tendency to produce class probability distributions that are pathological. The predictions tend to gravitate to values near zero or one, producing distributions that are "U"-shaped ([Kuhn and Johnson, 2013](https://scholar.google.com/scholar?hl=en&as_sdt=0%2C7&q=%22Applied+Predictive+Modeling%22&btnG=)). 

First, we create the task and a train-test-split:

```{r}
task = as_task_classif(cells, target = "class")
splits = partition(task)
cells_train = task$clone()$filter(splits$train)
cells_test = task$clone()$filter(splits$test)
```

Now, train the learner: 

```{r}
learner = lrn("classif.naive_bayes", predict_type = "prob")
learner$id <- "Uncalibrated Learner"
learner$train(cells_train)
```

And have it predict on the unseen test data:

```{r}
prediction = learner$predict(cells_test)
```

We will now asses the prediction quality, first measuring meodel performance with ROC AUC and Brier Score.

```{r}
measures = msrs(c("classif.auc", "classif.bbrier"))
prediction$score(measures)
```

Model performance alone does not tell us, if the model is well-calibrated. So let's assess if our model is well-calibrated. Spoiler: it is not. The first clue is the extremely U-shaped distribution of the probability scores:

```{r, message = FALSE}
hist(prediction$data$prob, breaks = (0:10)/10)
```

There are almost no cells with moderate probability estimates.

Then, assess if the model is calibrated with `calibrationplot()`. The calibration plot shows the relationship between the predicted probabilities and the true outcomes. The plot is divided into bins, and within each bin, the mean predicted probability and the mean observed outcome are calculated. The calibration plot can be smoothed by setting `smooth = TRUE`.

```{r}
library(ggplot2)
calibrationplot(list(learner), cells_test, smooth = TRUE)
```
The model is not well-calibrated. For predicted probabilities from 0 to 0.4, it seems to underestimate the true probability, for predicted probabilities higher than 0.4, it tends to overestimate the true probability.

We can also statistically test, if the model is well-calibrated. Possibilities are i.e. Spiegelhalter's Z-test and the HL-test. The p-values of these tests are implemented as calibration measures in `mlr3calibration`. A small p-value indicates poor calibration.

```{r}
print("Spiegelhalter's Z:")
prediction$score(spiegelhaltersz$new())
print("HL-test:")
prediction$score(hltest$new())
```
## Calibration

The good news is that we can do something about this. There are tools to "fix" the probability estimates so that they have better properties, such as falling along the diagonal lines in the diagnostic plots shown above. Different methods improve the predictions in different ways.

`mlr3calibration` implements three calibration strategies:

- Platt-Scaling `platt`: The original model is used to predict on the validation data set, returning predicted probabilities. We already know these probabilities are off, i.e., not well-calibrated. Then, we fit a simple logistic regression model, using the predicted probabilities from the original model as "x/feature" and the true labels as "y/target". In this way, we stack a second model on top of the first one: Original Model -> Probabilities -> Logistic Regression Model -> Calibrated Probabilities.

- Beta-Calibration `beta`: Similar idea to Platt, but instead of a plain logistic regression, we use a more flexible parametric model based on the Beta distribution. The predicted probabilities from the original model are transformed through a function that can represent asymmetric calibration curves (unlike Platt’s symmetric S-shaped curve). This allows better correction when probabilities are skewed. In short: Original Model → Probabilities → Beta Function Transformation → Calibrated Probabilities. The parameter `ab` or `abm` controls if an intercept should be included.

- Isotonic-Calibration `isotonic`: Here we don’t assume a parametric form at all. Instead, we fit a monotonic, piecewise constant function (via isotonic regression) that maps the original probabilities to calibrated ones. This is very flexible — it can fit almost any monotonic calibration curve, but it may overfit on small datasets. So: Original Model → Probabilities → Isotonic Regression (monotone step function) → Calibrated Probabilities.

You can also implement your own Calibrator and add it the the dictionary provided by `mlr3calibration`.

We will try out all calibration strategies for the Naive Bayes learner. To do so, we have to decide for a calibration strategy (when in the CV the calibrator should be trained). As we don't tune our Naive Bayes learner, `tunefirst` is not necessary. We decide to train one calibrator per fold, i.e. `perfold`.

If you want to compare the three calibration approaches with each other, it can be time-consuming to train each calibration approach using cross validation. In these cases, you can train an rr object, store the models and pass it to the PipeOpCalibrationPerFold. This means that the base learner has to be trained once in advance, and only the calibrators have to be trained in the PipeOpCalibrationPerFold using the holdout folds.

```{r Calibration via rr Object, results = "hide", warning=FALSE}

# Initialize base learner
learner <- lrn("classif.naive_bayes", predict_type = "prob")

# Initialize resampling strategy
rsmp <- rsmp("cv", folds = 5)

# Create rr object
rr <- resample(cells_train, learner, rsmp, store_models = TRUE)

# Initialze the calibrated learners
learner_calibrated_platt <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "platt"))
learner_calibrated_platt$id <- "platt"

learner_calibrated_beta <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "beta"))
learner_calibrated_beta$id <- "beta"

learner_calibrated_isotonic <- as_learner(calibPipeOp("calib.perfold",
                                      rr = rr,
                                      method = "isotonic"))
learner_calibrated_isotonic$id <- "isotonic"

# Train the calibrated learners
learner_calibrated_platt$train(cells_train)
learner_calibrated_beta$train(cells_train)
learner_calibrated_isotonic$train(cells_train)
```
Now, we evaluate the calibration:

```{r}
learner$train(cells_train)
learner$id <- "uncalibrated"
calibrationplot(list(learner, learner_calibrated_platt, learner_calibrated_beta, learner_calibrated_isotonic), cells_test, smooth = TRUE)
```
We can also compare calibration error:

```{r}
preds_uncal = learner$predict(cells_test)
preds_cal_platt = learner_calibrated_platt$predict(cells_test)
preds_cal_beta = learner_calibrated_beta$predict(cells_test)
preds_cal_isotonic = learner_calibrated_isotonic$predict(cells_test)

print("uncalibrated")
preds_uncal$score(ece$new())
print("platt")
preds_cal_platt$score(ece$new())
print("beta")
preds_cal_beta$score(ece$new())
print("isotonic")
preds_cal_isotonic$score(ece$new())
```
So we see that isotonic regression calibrates best, i.e. leads to best calibration error.
